
#Uploading Interest and Difficulty CSV Files
```{r}
I1 <- read.csv("interest.csv")
D1 <- read.csv("difficulty.csv")
```

#Converting Data Frames to Matrices
```{r, echo = FALSE}
library(dplyr) 
#Takes the matrix, gets rid of the names, and flips it so that the names are on the rows.
I2 <- select(I1, 2:7)   

#Turn the data frame into a matrix
I2 <- as.matrix(I2) 

#We assign the row names to be the names of the student
rownames(I2) <- I1$name 

D2 <- select(D1, 2:7) 
D2 <- as.matrix(D2)
rownames(D2) <- D1$name
```

#Generate a user-based similarity matrix based on cosine similarity using the ratings the class gave each unit. Matrix represents similarity of interests between students in the class.
```{r, echo = FALSE}
#Transpose the matrix so that multiplication occurs by students rather than units.

#We transpose the matrix letting units be in the rows and students be in the columns
I2 <- t(I2)

#NA can't be understood under cosine similarity so we substitute it to zero.
I2[is.na(I2)] <- 0
D2[is.na(D2)] <- 0 

#Generate the cosine similarity values for each pair of students.

install.packages("lsa") #You will need to install the lsa package to access the cosine command.
library("lsa")

#Generates the cosine similarity values as a new matrix.
I.SIM <- cosine(I2)

#Since each student will be most similar to themselves we want to remove that information
diag(I.SIM) <- NA
```

#Make a quick query to find out which students are most similar to you. 
```{r}

my.name <- "Juan Djuwadi" #My name
head(rownames(I.SIM[order(I.SIM[my.name,], decreasing = TRUE),]), n = 2) 

#Orders the column of the matrix corresponding to UNI according to similarity and returns top two UNI ids for the students who's interests are most similar.

#A: "Satra Ahmadzadeh" "Euki Sawada"  
```


#Creating a unit-based similarity matrix for difficulty to suggest units.
```{r}
#Collaborative Filter

#Find the cosine similarity for the difficulty 
D.SIM <- cosine(D2)

#People will be mot similar to themselves, so we set the diagonal values to NA. 
diag(D.SIM) <- NA
head(rownames(D.SIM[order(D.SIM["pred.dif",], decreasing = TRUE),]), n = 1) 
```

#Creating a composite measure from interest and difficulty, then constructing a similarity matrix using this measure
```{r}

library(tidyr)
library(lsa)

I3 <- gather(I1,unit,interest, 2:7) #We take the data of stated student interest and wrangle it. Columns from left to right: name, unit, interest. 

D3 <- gather(D1, name, difficulty) #Collapses, the dataframe to just the name and difficulty column.  

C1 <- data.frame(I3$name, I3$unit, I3$interest, D3$difficulty) 

names(C1) <- c("name", "unit", "interest", "difficulty") 
C1 <- filter(C1, difficulty > 0) #We filter the NA values and the values of 0.
C2 <- select(C1, "interest", "difficulty") #Only select the interest and difficulty columns.

#Run PCA
pc <- prcomp(C2)

#Extract PC1 loadings as new measure and attach to stid & unit

C3 <- data.frame(C1$name, C1$unit, pc$x) #We attach the translated values from the point cloud to a new data frame

C4 <- select(C3, C1.name, C1.unit, PC1) #We only take the first principal component and disregard the second one. 

#Remove int from unit label
C4$C1.unit <- gsub(".int", "", C4$C1.unit)

#Recreate unit by student matrix
C5 <- spread(C4, C1.name, PC1) #Spreads the principal components associated to the student name and the un 
row.names(C5) <- C5$C1.unit 
C5$C1.unit <- NULL  
C5 <- as.matrix(C5) 
C5 <- ifelse(is.na(C5), 0, C5) 
C5 <- t(C5) 

#Generate cosine similarity matrix for units
C.SIM <- cosine(C5)
diag(C.SIM) <- NA

#Search for most similar unit to "neural"
head(rownames(C.SIM[order(C.SIM["neural",], decreasing = TRUE),]), n = 1)

```