---
title: "collaborative-filter"
author: "Qingying Zhou"
date: "3/31/2019"
output: html_document
---

In HUDK4051 there are six units, we will use the ratings of these units that the class did earlier in terms of both interest and difficulty to produce individual suggestions about what unit to attempt next.

Start by uploading both the interest and difficulty csv files:
```{r}
## Merged the dumplicate rows and delete the missing rows
I1 <- read.csv("interest.csv")
I1 <- I1 %>% group_by(name) %>% summarise_all(funs(toString(unique(na.omit(.))))) 

D1 <- read.csv("difficulty.csv")
D1 <- D1 %>% group_by(name) %>% summarise_all(funs(toString(unique(na.omit(.)))))

```

Convert the data frames to matrices:
```{r, echo = FALSE}
I2 <- as.matrix(I1[-1]) %>% apply(2, as.numeric)
rownames(I2) <- I1$name
D2 <- as.matrix(D1[,-1]) %>% apply(2, as.numeric)
rownames(D2) <- D1$name
```

## Interest data: Generate a user-based similarity matrix based on cosine similarity using the ratings the class gave each unit. This matrix will represent the similarity of interests between students in the class.
```{r, echo = FALSE}

#Look at your data, missing values are coded with zero not NA. Investigate cosine similarity and decide whether your missing values should be coded with zeros or NAs. Explain your choice.

I2 <- t(I2) #Transpose the matrix so that multiplication occurs by students rather than units.
I2[is.na(I2)] <- 0

#Then we can generate the cosine similarity values for each pair of students

library(lsa)

I.SIM <- cosine(I2) #This command generates the cosine similarity values as a new matrix. Click on I.SIM in the Global Environment pane to see what it looks like.

diag(I.SIM) <- NA #Since each student will be most similar to themselves we want to remove that information
```

Now, we can make a quick query to find out which students are most similar to you 
```{r}
my.name <- "Qingying Zhou" #Input my name as it appears in the data set

head(rownames(I.SIM[order(I.SIM[my.name,], decreasing = TRUE),]), n = 3) #This code orders the column of the matrix corresponding to your UNI according to similarity and returns the top two UNI ids for the students who's interests are most similar to you
## "Megan Russell" "LingCai Yuan" "Xinke Song" 
## Adjusted cosine similarity?? since Megan's rating is so different than mine actually based on the scale difference
```

# centered cosine similarity (Pearson correlation)
```{r}
## Better intuition
## Take consideration of the raters' scaling differences
## 0 considered as average rating
I3 <- as.matrix(I1[-1]) %>% apply(2, as.numeric)
rownames(I3) <- I1$name
I3 <- t(I3)
scaled_I3 <- scale(I3, scale = FALSE)
scaled_I3[is.na(scaled_I3)] <- 0
adj_I.SIM <- cosine(scaled_I3)
diag(adj_I.SIM) <- NA
head(rownames(adj_I.SIM[order(adj_I.SIM[my.name,], decreasing = TRUE),]), n = 3)
#[1] "Megan Russell" "Yanyi Chen" "Melissa Mock" 
## Worse result even 

```




This is a basic collaborative filter! You have used information about interest across the class to generate an individual suggestion. Email one of your top matches, you may find them to be a good person to work with or ask questions during the semester.

Now create a unit-based, rather than student-based similarity matrix for difficulty. Then use your similarity matrix to provide a suggested next unit to a student who is looking for the unit that is most similar in terms of difficulty to the "prediction" unit. 
```{r}
D2[is.na(D2)] <- 0
D.SIM <- cosine(D2)
diag(D.SIM)
head(rownames(D.SIM[order(D.SIM["pred.dif",], decreasing = TRUE),]), n = 3)
##  "sna.dif"    "neural.dif"
```

Finally, educational settings have important differences to purely commercial settings such as film or product suggestions. In education we want people not to just follow their interests as they may simply choose things that are easy for them so they learn very little. To reduce this possibility with your collaborative filter create a composite measure from interest and difficulty, then construct a similarity matrix using this measure. (HINT: PCA). Once you have built the similarity matrix generate a suggestion for a student who has just completed the "prediction" unit.

```{r}
library(tidyr)
library(dplyr)
library(lsa)

## Integrate the int and diff columns together
I4 <- gather(I1, unit, interest, 2:7)
D3 <- gather(D1, name, difficulty)
C1 <- cbind(I4, D3[2])
C1$unit <- gsub(".int", "", C1$unit)
C2 <- C1[,3:4] %>% apply(2, as.numeric)
C2[is.na(C2)] <- 0 ## Filled with 0 for the missing data

## PCA
pc <- prcomp(C2)
C3 <- data.frame(C1$name, C1$unit, pc$x) 
C4 <- select(C3, C1.name, C1.unit, PC1) ## PC1: the most variability
C5 <- spread(C4, C1.unit, PC1)
C5 <- C5[,-1] %>% as.matrix()

## Cosine similarity matrix
C.SIM <- cosine(C5)
diag(C.SIM) <- NA

#the most similar units to "prediction"
head(rownames(C.SIM[order(C.SIM["pred",], decreasing = TRUE),]), n = 2)
## "sna"  "loop"
```

Once you have completed your collaborative filter you can return to it each time you are choosing a new unit to complete.
