---
title: "collaborative-filter"
author: "The real Andrea Maria Cornejo Hovispo"
date: "1/31/2019"
output: html_document
---

```{r}
install.packages("tidyr")
```



Start by uploading both the interest and difficulty csv files:
```{r}
i1 <- read.csv("interest.csv")
d1 <- read.csv("difficulty.csv")

```

Converting data frames to matrices to use matrix operations:
```{r, echo = FALSE}
#HINT: First you will need to remove the student ids as matrices can only contain one data type. You will then need to rename your row names with the student ids.
d2 <- d1[-c(25, 13, 1), ] 
i2 <- i1[-c(25, 13, 1), ] 

d3 <- d2[-1]
i3 <- i2[-1]

row.names(d3) <- d2$name
row.names(i3) <- i2$name

d3 <- as.matrix(d3)
i3 <- as.matrix(i3)

```

First, lets look at the interest data. We can generate a user-based similarity matrix based on cosine similarity using the ratings the class gave each unit. This matrix will represent the similarity of interests between students in the class.
```{r, echo = FALSE}
#First let's transpose the matrix so that multiplication occurs by students rather than units.

#Look at your data, missing values are coded with zero not NA. Investigate cosine similarity and decide whether your missing values should be coded with zeros or NAs. Explain your choice.

## Should probably be NA, otherwise zeros may be confused with numerical value assigned to  level of interest. However, if a large amount of observations contain NAs, these wouldn't be considered in calculating cosine similarity, which would be "wasting" data. So often this decision should be taken considering the trade off of losing data and having "biased" calculations and results. 
i3 <- t(i3)

#Then we can generate the cosine similarity values for each pair of students
#install.packages("lsa") #You will need to install the lsa package to access the cosine command.
#install.packages("lsa")
library(lsa)

I.SIM <- cosine(i3) #This command generates the cosine similarity values as a new matrix. Click on I.SIM in the Global Environment pane to see what it looks like. ( this is cosine simila matrix) 

diag(I.SIM) <- NA #Since each student will be most similar to themselves we want to remove that information (putting NA on diagonal)
```

Now, we can make a quick query to find out which students are most similar to you. 
```{r}
my.name <- "Jiaying Li" #Input your name as it appears in the data set - I am using a random person's name bc I wasn't here before 

head(rownames(I.SIM[order(I.SIM[my.name,], decreasing = TRUE),]), n = 2) #This code orders the column of the matrix corresponding to your UNI according to similarity and returns the top two UNI ids for the students who's interests are most similar to yours
```
#I wasnt included in the survey bc I started class late :( but if I would have been, my preferences would have been most similar to whoever Jiaying Li is. And the recommender indicates, Shuo-Yuan and Jiani Yani are most similar in interest!

This is a basic collaborative filter! You have used information about interest across the class to generate an individual suggestion. Email one of your top matches, you may find them to be a good person to work with or ask questions during the semester.

Now create a unit-based, rather than student-based similarity matrix for difficulty. Then use your similarity matrix to provide a suggested next unit to a student who is looking for the unit that is most similar in terms of difficulty to the "prediction" unit. 
```{r}

library(lsa)

I.SIM2 <- cosine(d3) #This command generates the cosine similarity values as a new matrix. Click on I.SIM in the Global Environment pane to see what it looks like.

diag(I.SIM2) 


```

Educational settings have important differences to purely commercial settings such as film or product suggestions. 
In education we want people not to just follow their interests as they may simply choose things that are easy for them so they learn very little. 
To reduce this possibility, we caan have the collaborative filter create a composite measure from interest and difficulty, then construct a similarity matrix using this measure using princial component analysis (PCA). 

Once the similarity matrix is built, we can use it to generate a suggestion for a student who has just completed the "prediction" unit.

```{r}
# first use gather to organize data
library(tidyr)
i4 <- gather(i1,unit,interest, 2:7)
d4 <- gather(d1, unit, difficulty, 2:7)

#gather() collects a set of column names and places them into a single “key” column. It also collects the cells of those #columns and places them into a single value column.

#now bring data together into one single dataset!
library(dplyr)
together <- data.frame(i4$name, i4$unit, i4$interest, d4$difficulty)
names(together) <- c("name", "unit", "interest", "difficulty")
together <- filter(together, difficulty > 0)  # filter rows where the diff > 0 - filter is like analog of select for columns
new <- select(together, "interest", "difficulty")

#Run PCA
pc <- prcomp(new)
#Extract PC1 loadings as new measure and attach to stid & unit
C3 <- data.frame(together$name, together$unit, pc$x)  ## x are transformed data points
C4 <- select(C3, together.name, together.unit, PC1)  
#Remove int from unit label
C4$together.unit <- gsub(".int", "", C4$together.unit)

#Recreate unit by student matrix
C5 <- spread(C4, together.name, PC1)
row.names(C5) <- C5$together.unit
C5$together.unit <- NULL
C5 <- as.matrix(C5)
C5 <- ifelse(is.na(C5), 0, C5)
C5 <- t(C5)      #now transposing 

#Generate cosine similarity matrix for units
C.SIM <- cosine(C5)
diag(C.SIM) <- NA  ## making all fields on diagonal = NA (bc a student will always be most similar to him/herself)

#Search for most similar unit to "neural"
head(rownames(C.SIM[order(C.SIM["neural",], decreasing = TRUE),]), n = 1)
head(rownames(C.SIM[order(C.SIM["viz",], decreasing = TRUE),]), n = 1)


```

Once the collaborative filter is complete, it can be used whenver a new unit from the course is selected, in order to figure out who would be best to collaborate with! 

In the example code, the neural networks module is most similar to the Social Network Analysis module, and Vizualization is most similar to NLP
