---
title: "collaborative-filter"
author: "Charles Lang"
date: "1/31/2019"
output: html_document
---
```{r}
#install.packages("lsa") #You will need to install the lsa package to access the cosine command.
library(lsa)
library(tidyverse)

```


In HUDK4051 there are six units, we will use your ratings of these units in terms of both interest and difficulty to produce individual suggestions about what unit to attempt next.

Start by uploading both the interest and difficulty csv files:
```{r}
interest <- read.csv('interest.csv')
difficulty <- read.csv('difficulty.csv')
```

```{r}
#Data cleaning

#Unite
interest2 <- unite(interest, col = 'name', .data$first, .data$last)

#Use distinct()/unique() to get distinct values - there are name duplicates
#Finding duplicates
interest2 %>% group_by(.data$name) %>% summarise(count = n()) %>% filter(`count` > 1)

#Dealing with duplicates - average duplicates
interest2 <- interest2 %>% group_by(.data$name) %>% summarise_all(.funs = mean, na.rm = TRUE)

```

We will be using matrix operations in this assignment, so convert your data frames to matrices:
```{r, echo = FALSE}
#HINT: First you will need to remove the student ids as matrices can only contain one data type. You will then need to rename your row names with the student ids.

#rownames() to preserve names
#rownames
interest2 <- column_to_rownames(interest2, var = "name")

interest2 <- as.matrix(interest2)

```

First, lets look at the interest data. We can generate a user-based similarity matrix based on cosine similarity using the ratings the class gave each unit. This matrix will represent the similarity of interests between students in the class.

```{r}
#First let's transpose the matrix so that multiplication occurs by students rather than units.
I2 <- t(interest2)
```

Look at your data, missing values are coded with zero not NA. Investigate cosine similarity and decide whether your missing values should be coded with zeros or NAs. Explain your choice.

Impressively there are no 0 values in the data, but for discussion's sake...

1. An NA value for cosine similarity within a vector prevents the calculation cosine similarity (returns NA). In theory, you cannot calculate the angular distance between two data points in an n-dimensional space (our metric of similarity when using cosine similarity) when you don't know where one (or both of the points are) in the spac,e because of a missing coordinate on one or more dimensions. In practice, it prevents the calculation of a dot product or vector magnitude.
2. A 0 value for missing data allows calculation of cosine similarity, but it is naturally only an estimate. For dimensions where the location of the data point in space is unknown, we get round the problem by just assuming it is 0. This could potentially seriously throw off the cosine similarity calculation - imagine two coordinates in a 2D space (1, 2) and (1, 3). Imagine that you didn't know the y-value of (1,3), and assumed it was (1,0). The real cosine similarity and estimated cosine similarity changes drastically. However, I assume this is mitigated if there are a lot more dimensions or dimensions with a smaller range, so it might be ok in these circumstances.

```{r, echo = FALSE}
#Then we can generate the cosine similarity values for each pair of students

I.SIM <- lsa::cosine(I2) #This command generates the cosine similarity values as a new matrix. Click on I.SIM in the Global Environment pane to see what it looks like.

diag(I.SIM) <- NA #Since each student will be most similar to themselves we want to remove that information
```

Now, we can make a quick query to find out which students are most similar to you. 
```{r}
my.name <- "XXXX" #Input your name as it appears in the data set

head(rownames(I.SIM[order(I.SIM[my.name,], decreasing = TRUE),]), n = 2) #This code orders the column of the matrix corresponding to your UNI according to similarity and returns the top two UNI ids for the students who's interests are most similar to yours
```

This is a basic collaborative filter! You have used information about interest across the class to generate an individual suggestion. Email one of your top matches, you may find them to be a good person to work with or ask questions during the semester.

Now create a unit-based, rather than student-based similarity matrix for difficulty. Then use your similarity matrix to provide a suggested next unit to a student who is looking for the unit that is most similar in terms of difficulty to the "prediction" unit. 
```{r}

```

Finally, educational settings have important differences to purely commercial settings such as film or product suggestions. In education we want people not to just follow their interests as they may simply choose things that are easy for them so they learn very little. To reduce this possibility with your collaborative filter create a composite measure from interest and difficulty, then construct a similarity matrix using this measure. (HINT: PCA). Once you have built the similarity matrix generate a suggestion for a student who has just completed the "prediction" unit.

```{r}


```

Once you have completed your collaborative filter you can return to it each time you are choosing a new unit to complete.
